Automated topic naming to support analysis of software maintenance activities

### 2010-08-18 icse/icse2011.tex
minor intro edits

\begin{abstract}
\todo[inline]{Topics and LDA sooner, Topics/concept location is a known
  research area and we're saying they don't do enough}
  Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.  
 Unfortunately the output of these tools is often difficult to distinguish and interpret as they are often so abstract.
 Thus to have a meaningful discussion about the topics of software development, we must be able to devise appropriate labels for extracted topics. 
 However, these approaches neither use domain-specific knowledge to improve results, nor contextualize those results for developers.   While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail. 
 This paper implements \emph{labelled topic extraction}, in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy. We focus on non-functional requirements related to software quality as a potential generalization, since there is some shared belief that these qualities apply broadly across many software systems and their development artifacts.  
 We evaluated our approach with an experimental study on two large-scale database projects, MySQL and MaxDB. We extracted topics using Latent Dirichlet Allocation (LDA) from the commit log comments of their version control systems (CVS and BitKeeper). Our results were generalizable across the two projects, showing that non-functional requirements were commonly discussed, and we identified topic trends over time. 	
 Our labelled topic extraction technique allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into software development activities.
\end{abstract}

@@

When a developer addresses a topic of development, such as a requirement, maintenance, bug fixes or other software engineering tasks, they often produce or modify multiple development artifacts.
%
These development artifacts include: source code, bug reports, revisions to source code, and even test cases. 
%
Topics that are addressed are sometimes not immediately resolved, some topics recur across time, yet others are only dealt with briefly.
%
The goal of concept location and topic analysis research is to semi-automatically recover some of these topics, and the related software artifacts.
% get neil's data mining point in there
The results of topic analysis and concept location are often very project specific, and not globally generalizable.
% get the interpretation aspect
As well much of the current research stops short of interpreting the purpose of the topic. That is, extracted topics are usually manually named and labelled by those who read the topic words.
% 
Thus in this paper, we propose a method of interpreting and labelling some topics, and their related artifacts, that is generalizable and relevant to multiple projects.

@@

% topics are numerous?
Deriving such topics is difficult. 
Raw topics from topic analysis are represented as unlabelled context-less word-lists or word distributions.
Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents. 
This manual approach is impractical when one has to handle more than one hundred different topics. 
To scale to meaningfully sized projects, we require automatic assistance to determine the topic label.

@@

There are many avenues of further investigation.  We want to investigate developer attitudes related to these labels: i.e., when we label a topic, was the developer expressing positive or negative qualities about that label?  It is difficult to map abstract qualities to particular messages. Is a ``quality'' discussion about more than just corrective maintenance? %In this study we take the position that any type of maintenance has to do with software quality.

>>>>

\begin{abstract}
\todo[inline]{Topics and LDA sooner, Topics/concept location is a known
  research area and we're saying they don't do enough}

  Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.  
 Unfortunately the output of these tools is often difficult to distinguish and interpret as they are often so abstract.
 Thus to have a meaningful discussion about the topics of software development, we must be able to devise appropriate labels for extracted topics. 
 However, these approaches neither use domain-specific knowledge to improve results, nor contextualize those results for developers.   While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail. 
 This paper implements \emph{labelled topic extraction}, in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy. We focus on non-functional requirements related to software quality as a potential generalization, since there is some shared belief that these qualities apply broadly across many software systems and their development artifacts.  
 We evaluated our approach with an experimental study on two large-scale database projects, MySQL and MaxDB. We extracted topics using Latent Dirichlet Allocation (LDA) from the commit log comments of their version control systems (CVS and BitKeeper). Our results were generalizable across the two projects, showing that non-functional requirements were commonly discussed, and we identified topic trends over time. 	
 Our labelled topic extraction technique allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into software development activities.
\end{abstract}

@@

Topics in software engineering practice refer to specific `threads' between project participants. These topics can include particular requirements, maintenance requests, bug fixes or other software engineering tasks. %NEIL: define topic specifically as akin to email 'thread'. It sounded too vague. Also changed from 'developers' to 'participants' because developers sounds too coding oriented. Our work could apply to testers, users, etc.
In addressing a topic of development participants often produce or modify multiple development artifacts.
%
These artifacts can include source code, bug reports, revisions to source code, and test cases. 
%
Topics can be in several potential states: they may not be immediately resolved, they may re-occur, yet others are only dealt with briefly  (e.g., a bug report that is closed).
% semi-automatically sounds too uncertain. I think the introduction can be more muscular and bold. Don't give the reviewer any room to start doubting.
The goal of concept location and topic analysis research is to recover these topics and the related software artifacts.
% get neil's data mining point in there
To date, the results of topic analysis and concept location are project specific, and not globally generalizable.
% get the interpretation aspect
Much of the current research stops short of interpreting the purpose of the topic. That is, extracted topics are usually manually named and labelled by those who read the topic words.
% 
In this paper we propose a method of interpreting and labelling topics, and their related artifacts, that is generalizable and relevant to multiple projects.

@@

% topics are numerous?
Deriving such topics is difficult. 
Raw topics from topic analysis are represented as unlabelled context-free word-lists or word distributions.
Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents. 
This manual approach is impractical when one has to handle hundreds of different topics; to scale to meaningfully sized projects, we require automatic assistance to determine the topic label.

@@

There are several avenues of further investigation.  We want to investigate developer attitudes related to these labels: i.e., when we label a topic, was the developer expressing positive or negative qualities about that label?  It is difficult to map abstract qualities to particular messages. Is a ``quality'' discussion about more than just corrective maintenance? %In this study we take the position that any type of maintenance has to do with software quality.
